---
title: "Statistical Inference Course Project Report"
author: "GP"
date: "12/21/2020"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE)
```

## Statistical Inference Course Project Report
Author: G. Purvinis

### Overview
This reports documents the final course project required by Coursera's "Statistical Inference" course. The project contains two parts. In Part 1, we are given an exponential distribution function as a population distribution, and asked to demonstrate the Central Limit Theorem (CLT). The CLT states that if we take n random samples from a population with mean mu and variance sigma^2, then the distribution of the sample mean will be approximately normal, no matter what the population distribution is, provided n is sufficiently large (typically n>30). Furthermore, the sample variance will equal the population variance divided by n. In part 2, basic inferential data analysis will be performed on the R Dataset package "ToothGrowth." A null hypothesis for the data will be formed and tested.
### Part 1: Simulations Investigating the Exponential Distribution Function

#### Sample Mean versus Theoretical Mean

#### Sample Variance versus Theoretical Variance

#### Distributions of the Sample Mean and Variance

#### Conclusion

### Part 2: Inferential Data Analyis Using "ToothGrowth" R Dataset package

#### The ToothGrowth dataset

#### Data Summary

#### Testing the data using Statistical Inference Techniques

#### Conclusion



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=TRUE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
