---
title: "Statistical Inference Course Project Report"
author: "GP"
date: "12/21/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Statistical Inference Course Project Report
Author: G. Purvinis

### Overview
This reports documents the final course project required of Coursera's "Statistical Inference" course. The project contains two tasks. In the first task (Part 1), we are given an exponential distribution function as a population distribution, and asked to demonstrate the Central Limit Theorem (CLT). The CLT states that if we take n random samples from a population with mean mu and variance sigma^2, then the distribution of the sample mean will be approximately normal, no matter what the population distribution is, provided n is sufficiently large (typically n>30). Furthermore, the sample variance will equal the population variance divided by n.
### Part 1: Simulations Investigating the Exponential Distribution Function

#### Sample Mean versus Theoretical Mean

#### Sample Variance versus Theoretical Variance

#### Distributions of the Sample Mean and Variance

#### Conclusion

### Part 2: Inferential Data Analyis Using "ToothGrowth" R Dataset package

#### The ToothGrowth dataset

#### Data Summary

#### Testing the data using Statistical Inference Techniques

#### Conclusion



```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=TRUE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
