---
title: "Statistical Inference Course Project Report, Part 1"
author: "GP"
date: "12/21/2020"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
library(knitr)
library(ggplot2)
library("cowplot")
knitr::opts_chunk$set(echo = TRUE)
```


### Overview
This reports documents the final course project required by Coursera's "Statistical Inference" course. The project contains two parts. In Part 1, we are given an exponential distribution function as a population distribution, and asked to demonstrate the Central Limit Theorem (CLT). The CLT states that if we take n random samples from a population with mean mu and variance sigma^2, then the distribution of the sample mean will be approximately normal, no matter what the population distribution is, provided n is sufficiently large (typically n>30). Furthermore, the sample variance will equal the population variance divided by n. 

The exponential distribution is a continuous distribution that is sometimes used to model the time that elapses before an event occurs.The function contains the parameter lambda, also called the rate parameter. For this exercise, lambda = 0.2 for all of the simulations.The mean of exponential distribution is 1/lambda and the standard deviation is also 1/lambda.

### Part 1: Simulations Investigating the Exponential Distribution Function
The plots below shows what the exponential distribution function looks like, both continuously (fig. 1A) and sampled.The R function rexp(n,rate) generates a vector of exponentially distributed random numbers. If we generate n=1000 exponentially distributed random numbers, then bin the numbers into a histogram, the distribution looks like figure 1B.

```{r ,echo=TRUE,fig.dim = c(6,3)}
lambda = 0.2
xs <- seq(0,40, by = .05)
dvalues <- dexp(xs,rate=lambda)
data <-data.frame(xs,dvalues) #data for continuous exp function
p1 <- ggplot(data,aes(x=xs,y=dvalues)) + geom_line(size = 2)+
   geom_area( fill="#69b3a2", alpha=0.4)+
   geom_vline(aes(xintercept = 5),color = "orange") +
      xlab("x") + ylab("0.2 exp(x/0.2)")+
   labs(caption="Fig.1A. Exponential Distribution Function \nwith rate lambda = 0.2",
        title ="Continuous") +
      annotate("text", x = 13, y = .12, 
              label = "mean = 5" , color = "orange",
              size=5 , angle=0, fontface="bold")
set.seed(12172020)  #set seed so result is reproducible
histdata <-data.frame(hx =rexp(1000,lambda))  #generate data for plot B
p2 <- ggplot(data=histdata,aes(x=hx))+
      geom_histogram(aes(y=..density..),fill="blue",alpha = 0.4,binwidth = 1)+
   geom_vline(aes(xintercept = 5),color = "orange")+
   xlab("x")+ ylab("distribution")+
   labs(caption="Fig.1B. Distribution of 1000 random \nexponentially distributed numbers",
        title ="Sampled")+
   annotate("text", x = 13, y = .12, 
              label = "sample \nmean ~ 5" , color = "orange",
              size=5 , angle=0, fontface="bold")
plot_grid(p1, p2, labels = c('A', 'B'), label_size = 12)

```


#### Sample Mean versus Theoretical Mean


#### Sample Variance versus Theoretical Variance

#### Distributions of the Sample Mean and Variance
Next we find the mean of 40 random exponentially distributed numbers, do this 1000 times, and plot the distribution of these means for the 1000 trials. The code below performs the function an plots the result. 

#### Conclusion

### Part 2: Inferential Data Analyis Using "ToothGrowth" R Dataset package

#### The ToothGrowth dataset

#### Data Summary

#### Testing the data using Statistical Inference Techniques

#### Conclusion

